Risk Perceptions for Wearable Devices

### 2015-03-05 tex-inputs/intro.tex
Some editing to intro, trimming down sentences and making it a little less pompous/formal

Wearable technologies, or ``wearables,'' are a \$700 million industry \cite{cmo} of electronically enhanced clothing items and accessories that interweaves technological interaction with everyday life. A top 25 market research company estimates that 52\% of technology consumers are aware of wearables and 33\% said they were likely to buy one~\cite{NPD}. With 20\% of the general population owning at least one wearable and 10\% using it daily~\cite{WearableStatNews}, wearables are transforming ubiquitous computing into a part of every day life. Forbes has named 2014 the ``Year of Wearable Technology \cite{Forbes}.''

The constantly captured data from these devices has many benefits, ranging from a more natural, human-centered interface experience to a healthier, fitness-data inspired lifestyle. There will likely be many more applications in the future which take advantage of such data. It is clear why wearable devices are becoming even more popular, especially as they have more capabilities and benefits over traditional devices.

Along with these benefits, wearable devices bring new potential privacy and security risks which expose users' activities without their awareness or consent. Fitbit allowed sex to be tracked as exercise while fitness profiles were public by default~\cite{Fitbit}, resulting in the inadvertent disclosure of sensitive information. Public discomfort toward facial recognition prevented the capability from being deployed to Google Glass \cite{GlassDetection}. Google Glass, the iconic wearable of its time, has since disappeared \cite{13_google_2015}. Most suspect that the reason for its disappearance was because it was not a shift in interest, but public concern over privacy issues \cite{14_dvorak_2014}. Some Glass wearers faced assault \cite{1_russell_2014, 15_mashable_2014, 16_gross_2014} from uncomfortable bystanders.

We have seen similar privacy \cite{kelley2013privacy, sadeh2009understanding, shklovski2014leakiness} and security issues \cite{enck2011study, felt2011survey} related to data capture with respect to smartphones. Mobile platforms have tried to address this by communicating data capture to users as the data is captured. However, many users are habituated to these notifications, because they see them all the time, often for things that they don't care about \cite{felt2012android}.

>>>>

Wearable technologies, or ``wearables,'' are a \$700 million industry \cite{cmo} that may see significant growth in the near future: one market research company estimates that 52\% of technology consumers are aware of wearables and 33\% said they were likely to buy one~\cite{NPD}. With 20\% of the general population owning at least one wearable and 10\% using it daily~\cite{WearableStatNews}, wearables are bringing ubiquitous computing to everyday life.

% Forbes has named 2014 the ``Year of Wearable Technology \cite{Forbes}.''

Wearable devices offer many compelling opportunities for benefiting users.  For instance, they could enable a more natural, human-centered interface experience and a healthier, fitness-data inspired lifestyle. There will likely be many more applications in the future which take advantage of data captured by wearable devices.

However, wearable devices also bring new potential privacy and security risks that could expose users' activities without their awareness or consent. Even though wearable technology is currently in early stages, we have already seen instances of these risks. For instance, Fitbit's default privacy settings inadvertently exposed information about some of their users' sexual activity~\cite{Fitbit}, resulting in the inadvertent disclosure of sensitive information. Public discomfort toward facial recognition caused Google to prohibit Google Glass apps from using facial recognition~\cite{GlassDetection}, and public concerns about privacy may have partly contributed to Google's withdrawal of Glass~\cite{14_dvorak_2014,1_russell_2014, 15_mashable_2014, 16_gross_2014}.

We have seen similar privacy \cite{kelley2013privacy, sadeh2009understanding, shklovski2014leakiness} and security issues \cite{enck2011study, felt2011survey} related to data capture on smartphones. Mobile platforms have tried to address the risks by communicating data capture to users as the data is captured. However, many users are habituated to these notifications, because they see them all the time, often for things that they don't care about \cite{felt2012android}.
